---
title: "HW5_hwenjun"
author: "Wenjun Han"
date: "10/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
```

## Problem 3

Get and clean the following data on education from the World Bank. 
```{r clean data}
# Read the csv data from the dovwloaded csv files
edstat_raw <- read.csv(file = 'EdStatsData.csv')

# Clean the data through tidyverse
# First we restack the data into a right format
edstat <-edstat_raw %>%
                      gather(key = "Time",value = "value", X1970:X2100)
# We found that the data has an X variable which not exist in orginal data
edstat = subset(edstat, select = -X)

#Change name of the variables
names(edstat)[1] <- "Country.Name"
head(edstat)

# Since we found that there are many missing data
# We remove the missing data from our table
edstat <- drop_na(edstat)
head(edstat)
dim(edstat)
```

There are 57650450 data points in complete dataset, and 5082201 data points we have in cleaned dataset. We chose country Australia
and Hungary. 

```{r Summary table of indicators}
# Choose country Australia and Hungary
library(dplyr)
aus <- edstat[edstat$Country.Name == "Australia",]
head(aus)
hun <- edstat[edstat$Country.Name == "Hungary",]
head(hun)

# Create a summary table of indicators for comparison
summary(aus)
summary(hun)
by_aus <- aus %>% group_by(Indicator.Code) %>% summarise(mean = mean(value), n = n())
head(by_aus)
by_hun <- hun %>% group_by(Indicator.Code) %>% summarise(mean = mean(value), n = n())
head(by_hun)
```

The summary of indicators of Australia data and Hungary data is shown as above tables.


## Problem 4

Using base plotting functions to create a single figure as similar to the example. Use simple linear regression method for plotting data.

```{r base plot}
library(MASS)
library(faraway)
library(olsrr)
library(sur)

# Plot country Australia
# Fit data with linear regression model
lmfit_aus <- lm(aus$value~aus$Indicator.Code)

# Plot 1
par(mfcol=c(2,3))
plot(fitted(lmfit_aus),residuals(lmfit_aus),xlab = "Predicted Value",ylab = "Residuals")
abline(h=0)

# Plot 2
plot(fitted(lmfit_aus),studres(lmfit_aus),xlab = "Predicted Value",ylab = "Rstudent")

# Plot 3
plot(leverage(lmfit_aus),studres(lmfit_aus),xlab = "Leverage",ylab = "Rstudent")

# Plot 4
qqnorm(residuals(lmfit_aus),ylab = "Residuals")

# Plot 5
plot(fitted(lmfit_aus),weighted.residuals(lmfit_aus, drop0 = TRUE),xlab = "Predicted Value",ylab = "Weight")

# Plot 6
plot(cooks.distance(lmfit_aus),xlab = "Observation",ylab = "Cook's D")


# Plot country Hungary
lmfit_hun <- lm(hun$value~hun$Indicator.Code)

par(mfcol=c(2,3))
plot(fitted(lmfit_hun),residuals(lmfit_hun),xlab = "Predicted Value",ylab = "Residuals")
abline(h=0)

# Plot 2
plot(fitted(lmfit_hun),studres(lmfit_hun),xlab = "Predicted Value",ylab = "Rstudent")

# Plot 3
plot(leverage(lmfit_hun),studres(lmfit_hun),xlab = "Leverage",ylab = "Rstudent")

# Plot 4
qqnorm(residuals(lmfit_hun),ylab = "Residuals")

# Plot 5
plot(fitted(lmfit_hun),weighted.residuals(lmfit_hun, drop0 = TRUE),xlab = "Predicted Value",ylab = "Weight")

# Plot 6
plot(cooks.distance(lmfit_hun),xlab = "Observation",ylab = "Cook's D")

```

## Problem 5

Recreate the plot in the last problem using ggplot2 functions.

```{r gg plot}
library(tidyverse)
library(ggplot2)
library(ggpubr)

# Plot country Australia

# Plot 1

lmfit_ausf <- fortify(lmfit_aus)
a1 <- ggplot(lmfit_ausf, aes(x = .fitted, y = .resid)) + geom_point()

# Plot 2
b1 <- ggplot(lmfit_ausf, aes(x = .fitted, y = .stdresid)) + geom_point()

# Plot 3
c1 <- ggplot(lmfit_ausf, aes(x = leverage(lmfit_aus), y = .stdresid)) + geom_point()

# Plot 4
d1 <- ggplot(lmfit_ausf, aes(sample=.resid))+stat_qq()

# Plot 5
e1 <- ggplot(lmfit_ausf, aes(x = .fitted, y = weighted.residuals(lmfit_aus, drop0 = TRUE))) + geom_point()

# Plot 6
f1 <- ggplot(lmfit_ausf, aes(x= c(1:28083), y=.cooksd)) + geom_point()

ggarrange(a1, b1, c1, d1, e1, f1,
          ncol = 3, nrow = 2)



# Plot country Hungary
lmfit_hunf <- fortify(lmfit_hun)
a2 <- ggplot(lmfit_hunf, aes(x = .fitted, y = .resid)) + geom_point()

# Plot 2
b2 <- ggplot(lmfit_hunf, aes(x = .fitted, y = .stdresid)) + geom_point()

# Plot 3
c2 <- ggplot(lmfit_hunf, aes(x = leverage(lmfit_hun), y = .stdresid)) + geom_point()

# Plot 4
d2 <- ggplot(lmfit_hunf, aes(sample=.resid))+stat_qq()

# Plot 5
e2 <- ggplot(lmfit_hunf, aes(x = .fitted, y = weighted.residuals(lmfit_hun, drop0 = TRUE))) + geom_point()

# Plot 6
f2 <- ggplot(lmfit_hunf, aes(x= c(1:33046), y=.cooksd)) + geom_point()

ggarrange(a2, b2, c2, d2, e2, f2,
          ncol = 3, nrow = 2)

```









